---
title: "Lab 03 — Regression for Description and Prediction"
subtitle: "Sessions 6–7: Regression, Parts 1–2"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---


## Learning goals

You will use simple linear regression for **description and prediction**, interpret coefficients, calculate fitted values, and examine residuals.

## 1. Setup

```{r}
library(tidyverse)
world <- read_csv("data/country_indicators.csv")
```

## 2. A regression question

> Can GDP per capita predict life expectancy?

```{r}
m1 <- lm(life_expectancy ~ log(gdp_per_capita), data = world)
summary(m1)
```

Write the fitted equation in the form

\[
\widehat{Y} = \widehat{\beta}_0 + \widehat{\beta}_1 X.
\]

What does the slope mean in substantive terms?

## 3. Add the fitted line

```{r}
ggplot(world, aes(x = log(gdp_per_capita), y = life_expectancy)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Predicting life expectancy from economic development",
    x = "Log GDP per capita",
    y = "Life expectancy"
  )
```

## 4. Prediction

Create a prediction for a hypothetical country with GDP per capita of 20,000.

```{r}
predict(
  m1,
  newdata = data.frame(gdp_per_capita = 20000)
)
```

Now predict for 5,000 and 50,000.

## 5. Residuals

```{r}
world_with_fit <- world |>
  mutate(
    predicted = predict(m1),
    residual = residuals(m1)
  )

world_with_fit |>
  select(country, life_expectancy, predicted, residual) |>
  arrange(desc(abs(residual))) |>
  slice_head(n = 8)
```

What does a positive residual mean? What does a negative residual mean?

## 6. Description is not causation

Does the slope show that increasing GDP would *cause* life expectancy to increase? Explain.

## DELIVERABLE — Regression Memo

Submit a **Regression Memo** containing:

- the research question;
- the estimated intercept and slope;
- one fitted regression figure;
- a prediction for a substantively meaningful value of X;
- one country with a large residual and an interpretation;
- **150–220 words** explaining what the model can describe/predict and what it cannot establish causally.

Render and submit the HTML.
