---
title: "Lab 03 — Regression for Description and Prediction"
subtitle: "Sessions 6–7: Regression, Parts 1–2"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---

## Learning goals

You will fit one line, read the slope as an association, and separate a fitted value from an effect. The file is the candidate-face study in QSS 4.2. Subjects rated how competent unfamiliar candidates looked. `d.comp` is the Democrat's competence rating. The Democratic margin in the vote is built from the vote counts.

A face rating that predicts vote share is not a policy lever.

## 1. Setup

```{r}
library(tidyverse)
face <- read_csv("data/face.csv") %>%
  mutate(
    d.share = d.votes / (d.votes + r.votes),
    r.share = r.votes / (d.votes + r.votes),
    diff.share = d.share - r.share
  )
fit <- lm(diff.share ~ d.comp, data = face)
coef(fit)
```

The slope is about 0.66. In this sample, a one-point higher Democratic competence rating goes with about a 0.66 higher Democratic margin in vote share. The correlation is about 0.43. That is in-sample association.

```{r}
ggplot(face, aes(x = d.comp, y = diff.share)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Democratic competence rating",
    y = "Democratic margin in vote share",
    title = "Faces and vote share"
  )
```

The line is a summary of these candidates. A fitted value is the margin the line predicts at a rating. It is not the effect of changing a candidate's face.

## Source

`face.csv` is taken unchanged from [kosukeimai/qss](https://github.com/kosukeimai/qss) `PREDICTION/face.csv`. Sessions 6–7 in [mir-prob-stats-QQS](https://github.com/iedaejin/mir-prob-stats-QQS). See `data/SOURCE.txt`.
