---
title: "Lab 06 — Reversion to the Mean and Correlation vs Causation"
subtitle: "Sessions 13–14: Reversion to the Mean; Correlation and Causation"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---

## Learning goals

You will see reversion to the mean in two real elections, and you will keep that pattern separate from a causal claim. The files are the 2008 and 2012 state presidential results used in QSS 4.2.4. Obama's state vote share is very stable across the two years. The slope of the 2012 z-score on the 2008 z-score equals the correlation, about 0.98. Because that slope is below 1, a state that was extreme in 2008 is predicted to be slightly less extreme in 2012.

## 1. Setup

```{r}
library(tidyverse)
pres08 <- read_csv("data/pres08.csv")
pres12 <- read_csv("data/pres12.csv")
pres <- inner_join(
  pres08 %>% select(state, Obama08 = Obama),
  pres12 %>% select(state, Obama12 = Obama),
  by = "state"
) %>%
  mutate(
    z08 = as.numeric(scale(Obama08)),
    z12 = as.numeric(scale(Obama12))
  )
coef(lm(z12 ~ z08, data = pres))
```

## 2. The states that were low in 2008

```{r}
pres %>%
  filter(z08 < quantile(z08, 0.25)) %>%
  summarise(
    n = n(),
    mean_z08 = mean(z08),
    mean_z12 = mean(z12),
    share_higher = mean(z12 > z08)
  )
```

Those states were about 1.18 standard deviations below the 2008 mean. In 2012 they were about 1.13 below the mean, and a majority of them moved up. Nothing in this comparison is a program that treated the low states. Part of an extreme 2008 share was particular to that year. A ministry that credits a later program for the improvement is ignoring that rival story.

A single cross-section, one year of countries, cannot show this. Reversion needs the same units measured twice.

## Source

`pres08.csv` and `pres12.csv` are taken unchanged from [kosukeimai/qss](https://github.com/kosukeimai/qss) `PREDICTION/`. Session 13 in [mir-prob-stats-QQS](https://github.com/iedaejin/mir-prob-stats-QQS). See `data/SOURCE.txt`.
