---
title: "Lab 06 — Reversion to the Mean and Correlation vs Causation"
subtitle: "Sessions 13–14: Reversion to the Mean; Correlation and Causation"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---


## Learning goals

You will see reversion to the mean in simulated data and then revisit the distinction between correlation and causation.

## 1. Setup

```{r}
library(tidyverse)
world <- read_csv("data/country_indicators.csv")
```

## 2. Reversion to the mean

Suppose public approval of an international organization is measured twice. Each observed score contains a stable component plus temporary noise.

```{r}
set.seed(321)

n <- 500
stable_support <- rnorm(n, mean = 50, sd = 10)

sim <- tibble(
  wave1 = stable_support + rnorm(n, 0, 8),
  wave2 = stable_support + rnorm(n, 0, 8)
)
```

Select respondents with unusually high values at Wave 1.

```{r}
extreme_high <- sim |>
  filter(wave1 >= quantile(wave1, 0.90))

extreme_high |>
  summarise(
    mean_wave1 = mean(wave1),
    mean_wave2 = mean(wave2)
  )
```

Why is the second mean usually less extreme?

## 3. Visualize it

```{r}
ggplot(sim, aes(x = wave1, y = wave2)) +
  geom_point(alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  labs(
    title = "Repeated measurements and reversion to the mean",
    x = "Wave 1",
    y = "Wave 2"
  )
```

## 4. Correlation and causation

Now return to country data.

```{r}
m_naive <- lm(life_expectancy ~ democracy_score, data = world)
summary(m_naive)
```

Add economic development:

```{r}
m_adjusted <- lm(
  life_expectancy ~ democracy_score + log(gdp_per_capita),
  data = world
)
summary(m_adjusted)
```

Compare the democracy coefficient across the two models.

## 5. Causal diagnosis

Draw or describe the causal structure you think could produce these associations.

For example:

\[
GDP \rightarrow Democracy
\]

\[
GDP \rightarrow Life\ Expectancy
\]

\[
Democracy \rightarrow Life\ Expectancy\ ?
\]

Why can adjustment change a coefficient?

## DELIVERABLE — Causal Diagnosis

Submit a **Causal Diagnosis** containing:

- the democracy coefficient from the naive model;
- the democracy coefficient from the adjusted model;
- a simple causal diagram drawn by hand and inserted as an image **or** written using arrows in the notebook;
- **180–250 words** explaining:
  - why the coefficient changed;
  - what confounding means in this example;
  - why neither regression alone proves causality;
  - one type of evidence that would strengthen a causal claim.

Submit the rendered HTML.
