---
title: "Lab 01 — Causal Questions and Describing Data"
subtitle: "Sessions 2–3: Causation; Describing a Single Variable"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---

## Learning goals

You will separate a description from an effect question, then summarise one variable. The file is the résumé audit study in QSS 2.1–2.3. Each row is a fictitious résumé. `call` is 1 if the employer called back. `race` records whether the name on the résumé was assigned to sound Black or White. The name was randomized, so a gap in callback rates estimates discrimination in callback, not a difference in the résumé's qualifications.

## 1. Setup

```{r}
library(tidyverse)
resume <- read_csv("data/resume.csv")
glimpse(resume)
```

## 2. Two questions, two jobs

- Descriptive: what share of these résumés received a callback?
- Causal: what is the effect of a Black-sounding name, relative to a White-sounding name, on the callback?

The mean of a 0/1 variable is a proportion.

```{r}
resume %>%
  summarise(callback = mean(call), n = n())
```

```{r}
resume %>%
  group_by(race) %>%
  summarise(callback = mean(call), n = n(), .groups = "drop")
```

White-sounding names were called back about 9.6 percent of the time. Black-sounding names, about 6.4 percent. The difference is about 3.2 percentage points.

Only one name was placed on each résumé. The callback under the other name is the counterfactual, and it is not in the file. The randomized name makes the *average* gap an estimate of the average effect. It does not reveal the effect for any single résumé.

## Source

`resume.csv` is taken unchanged from [kosukeimai/qss](https://github.com/kosukeimai/qss) `CAUSALITY/resume.csv` (Imai and Webb Williams). Session 2 in [mir-prob-stats-QQS](https://github.com/iedaejin/mir-prob-stats-QQS). See `data/SOURCE.txt`.
