---
title: "Lab 01 — Causal Questions and Describing Data"
subtitle: "Sessions 2–3: Causation; Describing a Single Variable"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---


## Learning goals

This lab follows the course sequence from **causal questions** to **describing a single variable**.

You will learn to distinguish a causal question from a descriptive question, identify an outcome and an explanatory variable, and summarize one variable using numerical and graphical tools.

## 1. Setup

```{r}
library(tidyverse)
world <- read_csv("data/country_indicators.csv")
```

## 2. From a substantive question to variables

Consider the question:

> Does economic development increase democratic governance?

Write down:

- **Outcome (Y):**
- **Explanatory variable / treatment (X):**
- **Two possible confounders:**
- **Why is this a causal rather than purely descriptive question?**

Do not run any model yet.

## 3. Describe one variable first

Start with `democracy_score`.

```{r}
world |>
  summarise(
    n = n(),
    mean = mean(democracy_score),
    median = median(democracy_score),
    sd = sd(democracy_score),
    min = min(democracy_score),
    max = max(democracy_score)
  )
```

What does each statistic tell you?

## 4. Look at the distribution

```{r}
ggplot(world, aes(x = democracy_score)) +
  geom_histogram(bins = 10) +
  labs(
    title = "Democracy scores across countries",
    x = "Democracy score",
    y = "Count"
  )
```

Is the distribution symmetric, skewed, clustered, or unusual?

## 5. Compare groups descriptively

```{r}
world |>
  group_by(region) |>
  summarise(
    mean_democracy = mean(democracy_score),
    median_democracy = median(democracy_score),
    n = n()
  ) |>
  arrange(desc(mean_democracy))
```

Why should we be cautious about interpreting these regional differences causally?

## 6. Choose your own variable

Choose one of:

- `gdp_per_capita`
- `life_expectancy`
- `trade_openness`
- `military_spending_pct_gdp`
- `un_voting_alignment`

Create:

1. a numerical summary;
2. a histogram;
3. a short substantive interpretation.

## DELIVERABLE — Country Data Brief

Submit a **one-page Country Data Brief** inside your rendered HTML.

It must contain:

- a clear descriptive research question;
- **one figure**;
- **at least three descriptive statistics**;
- **120–180 words** interpreting the distribution;
- one sentence explaining what **cannot** be inferred causally from the description.

Your brief should be understandable to a policy reader who has not seen your R code.
