---
title: "Lab 08 — Controlling for Confounders"
subtitle: "Sessions 18–19: Controlling for Confounders, Parts 1–2"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---


## Learning goals

You will analyze observational data, see how confounding can change an estimate, and compare unadjusted and adjusted regression models.

## 1. Setup

```{r}
library(tidyverse)
aid <- read_csv("data/aid_observational.csv")
```

The data are synthetic and represent country-level foreign-aid allocations.

Question:

> Does receiving more foreign aid increase economic growth?

## 2. Start with the raw association

```{r}
ggplot(aid, aes(x = aid_per_capita, y = economic_growth_pct)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Foreign aid and economic growth",
    x = "Aid per capita",
    y = "Economic growth (%)"
  )
```

## 3. Naive model

```{r}
m1 <- lm(economic_growth_pct ~ aid_per_capita, data = aid)
summary(m1)
```

Interpret the aid coefficient.

## 4. Think about selection into treatment

Countries do not receive foreign aid randomly.

In this teaching dataset, aid allocation is related to:

- baseline need;
- conflict intensity;
- institutional quality.

Explain why these variables might affect both aid receipt and subsequent economic growth.

## 5. Adjust for confounders

```{r}
m2 <- lm(
  economic_growth_pct ~ aid_per_capita +
    baseline_need +
    conflict_intensity +
    institutional_quality,
  data = aid
)

summary(m2)
```

Compare the aid coefficient in `m1` and `m2`.

## 6. Model comparison

```{r}
coef(m1)["aid_per_capita"]
coef(m2)["aid_per_capita"]
```

Does adjustment change sign, size, or uncertainty?

## 7. What assumptions remain?

Regression adjustment does not automatically solve confounding.

Name at least two assumptions or concerns that remain.

## DELIVERABLE — Confounding Memo

Submit a **Model Comparison Memo** containing:

- one scatterplot;
- the aid coefficient from the naive model;
- the aid coefficient from the adjusted model;
- a short table or clearly formatted comparison;
- **200–280 words** explaining:
  - why treatment assignment is not random;
  - which variables are potential confounders;
  - why the estimates changed;
  - whether you are willing to make a causal claim, and under what assumptions.

Submit the rendered HTML.
