---
title: "Lab 09 — Mechanisms"
subtitle: "Session 20: Assessing Mechanisms"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---


## Learning goals

You will distinguish a treatment effect from a proposed causal mechanism and use regressions to examine evidence consistent with a mechanism.

## 1. Setup

```{r}
library(tidyverse)
mech <- read_csv("data/mechanism_diplomatic_contact.csv")
```

The synthetic study asks whether diplomatic contact increases public support for cooperation.

Proposed mechanism:

\[
Diplomatic\ Contact \rightarrow Trust \rightarrow Support\ for\ Cooperation
\]

## 2. Total association between treatment and outcome

```{r}
m_total <- lm(
  support_for_cooperation ~ diplomatic_contact,
  data = mech
)

summary(m_total)
```

## 3. Does treatment affect the proposed mechanism?

```{r}
m_mech <- lm(
  trust_in_foreign_partner ~ diplomatic_contact,
  data = mech
)

summary(m_mech)
```

What would we expect if trust is part of the mechanism?

## 4. Is the mechanism associated with the outcome?

```{r}
m_out <- lm(
  support_for_cooperation ~ trust_in_foreign_partner,
  data = mech
)

summary(m_out)
```

## 5. Add both variables

```{r}
m_both <- lm(
  support_for_cooperation ~ diplomatic_contact +
    trust_in_foreign_partner,
  data = mech
)

summary(m_both)
```

Compare the diplomatic-contact coefficient in `m_total` and `m_both`.

## 6. Important caution

A smaller treatment coefficient after including the mediator is **not, by itself, proof of mediation**.

Discuss at least one reason why causal mechanism claims require stronger assumptions or design.

## DELIVERABLE — Mechanism Brief

Submit a **Causal Mechanism Brief** with:

- the proposed mechanism written as an arrow diagram;
- the treatment → outcome coefficient;
- the treatment → trust coefficient;
- the trust → outcome association;
- **200–280 words** answering:
  - Is the observed evidence consistent with the proposed mechanism?
  - What evidence is still missing?
  - What additional design or data would make the mechanism claim more convincing?

End with one sentence that clearly distinguishes **evidence of an effect** from **evidence about how the effect occurs**.

Submit the rendered HTML.
