---
title: "Lab 04 — Probability Through Simulation"
subtitle: "Sessions 9–10: Probability Theory, Parts 1–2"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---


## Learning goals

You will use simulation to understand probability, conditional probability, random variables, and expected values.

## 1. Setup

```{r}
library(tidyverse)
survey <- read_csv("data/survey_attitudes.csv")
```

## 2. Probability as long-run frequency

Imagine that 30% of voters support a diplomatic initiative. Simulate 1,000 voters.

```{r}
support <- rbinom(1000, size = 1, prob = 0.30)
mean(support)
```

Repeat the simulation. Why does the answer change slightly?

## 3. Sample size and stability

```{r}
mean(rbinom(20, size = 1, prob = 0.30))
mean(rbinom(200, size = 1, prob = 0.30))
mean(rbinom(20000, size = 1, prob = 0.30))
```

What happens as the number of trials increases?

## 4. Conditional probability in an IR context

Define “high support for international cooperation” as a score of 7 or more.

```{r}
survey <- survey |>
  mutate(high_support = support_international_cooperation >= 7)
```

Estimate:

```{r}
mean(survey$high_support)
```

Now estimate the probability among postgraduate respondents.

```{r}
survey |>
  filter(education == "Postgraduate") |>
  summarise(prob_high_support = mean(high_support))
```

How is this different from the unconditional probability?

## 5. A random variable and expectation

Suppose the number of diplomatic agreements signed in a year follows a binomial model with 10 opportunities and probability 0.4.

```{r}
agreements <- rbinom(10000, size = 10, prob = 0.4)

mean(agreements)
var(agreements)
```

What is the theoretical expected value \(E[X]\) for a Binomial(10, 0.4)?

## 6. Visualize the random variable

```{r}
tibble(agreements = agreements) |>
  ggplot(aes(x = agreements)) +
  geom_histogram(binwidth = 1, boundary = -0.5) +
  labs(
    title = "Simulated number of diplomatic agreements",
    x = "Agreements",
    y = "Frequency"
  )
```

## DELIVERABLE — Probability Simulation Note

Create a short simulation answering:

> How does sample size affect the stability of an estimated probability?

Your deliverable must include:

- estimates from **three different sample sizes**;
- one figure;
- the theoretical probability being estimated;
- **120–180 words** explaining why estimates vary and why larger samples are more stable.

Submit the rendered HTML.
