---
title: "Lab 04 — Probability Through Simulation"
subtitle: "Sessions 9–10: Probability, Parts 1–2"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---

## Learning goals

You will treat a probability as a long-run share, then compute a conditional probability on a real voter file. The simulation is the QSS 6.1 idea. The file is the Florida voter sample in QSS 6.2 (`FLVoters.csv`, 10,000 records, with surname, county, age, gender, and race).

## 1. A long-run share

Suppose 30 percent of draws are "yes." A short run jumps around. A long run sits near 0.30.

```{r}
library(tidyverse)
set.seed(4)
mean(rbinom(20, size = 1, prob = 0.30))
mean(rbinom(5000, size = 1, prob = 0.30))
```

## 2. Conditional probability in the voter file

```{r}
voters <- read_csv("data/FLVoters.csv") %>%
  filter(!is.na(race), !is.na(gender))
nrow(voters)
```

The filter drops the records with no race or no gender. 9,118 of the 10,000 remain, and every share below is a share of those.

```{r}
mean(voters$race == "black")
mean(voters$race[voters$gender == "f"] == "black")
```

About 13 percent of these voters are recorded as Black. Among women, the share is about 14 percent. `P(Black | woman)` is the share of women who are Black. It is not the share of all voters who are Black, and it is not the share who are Black women.

## Source

`FLVoters.csv` is taken unchanged from [kosukeimai/qss](https://github.com/kosukeimai/qss) `PROBABILITY/FLVoters.csv`. Sessions 9–10 in [mir-prob-stats-QQS](https://github.com/iedaejin/mir-prob-stats-QQS). See `data/SOURCE.txt`.
