---
title: "Lab 00 — Welcome to R"
subtitle: "Session 1: Introduction"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---

## Learning goals

By the end of this lab you should be able to:

- open and run an R code chunk;
- import a `.csv` file;
- name the unit of each row;
- calculate a simple summary;
- make a basic plot with `ggplot2`.

The file is the UN world-population series used in QSS Chapter 1. `world.pop` is recorded in thousands of people. One row is a year, not a country.

## 1. Setup

```{r}
library(tidyverse)
un <- read_csv("data/UNpop.csv")
un <- un %>% mutate(pop_millions = world.pop / 1000)
un
```

## 2. What is a row?

```{r}
glimpse(un)
nrow(un)
```

Seven rows. The series runs from 1950 to 2010, once a decade.

## 3. A summary and a plot

```{r}
un %>%
  summarise(
    mean_millions = mean(pop_millions),
    median_millions = median(pop_millions),
    first = min(pop_millions),
    last = max(pop_millions)
  )
```

```{r}
ggplot(un, aes(x = year, y = pop_millions)) +
  geom_line() +
  geom_point() +
  labs(
    x = "Year",
    y = "World population (millions)",
    title = "UN world population, 1950–2010"
  )
```

The line rises in every decade. That is a description of this series. It does not say which policy caused the increase.

## Source

`UNpop.csv` is taken unchanged from Kosuke Imai and Nora Webb Williams, *Quantitative Social Science* ([kosukeimai/qss](https://github.com/kosukeimai/qss), file `INTRO/UNpop.csv`), the dataset linked for Session 1 in [mir-prob-stats-QQS](https://github.com/iedaejin/mir-prob-stats-QQS). See `data/SOURCE.txt`.
