---
title: "Lab 00 — Welcome to R"
subtitle: "Session 1: Introduction"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---


## Learning goals

By the end of this lab you should be able to:

- open and run an R code chunk;
- import a `.csv` file;
- inspect a data frame;
- calculate a simple summary;
- make a basic plot with `ggplot2`;
- render a Quarto document to HTML.

## 1. Setup

```{r}
library(tidyverse)
```

## 2. Import the data

The dataset contains synthetic indicators for 50 countries. These are **teaching data**, not real country estimates.

```{r}
world <- read_csv("data/country_indicators.csv")
```

Inspect the first rows:

```{r}
head(world)
```

How many countries and variables are in the dataset?

```{r}
dim(world)
```

## 3. Ask R a simple question

What is the average democracy score?

```{r}
mean(world$democracy_score)
```

Now calculate the median democracy score.

```{r}
# Your code here
```

## 4. Make your first figure

```{r}
ggplot(world, aes(x = democracy_score)) +
  geom_histogram(bins = 10) +
  labs(
    title = "Distribution of democracy scores",
    x = "Democracy score",
    y = "Number of countries"
  )
```

## 5. Filter the data

Choose one region and inspect only those countries.

```{r}
world |>
  filter(region == "Europe")
```

Change `"Europe"` to another region.

## 6. A first substantive question

Which five countries in this teaching dataset have the highest GDP per capita?

```{r}
world |>
  arrange(desc(gdp_per_capita)) |>
  select(country, gdp_per_capita) |>
  slice_head(n = 5)
```

## DELIVERABLE — Your first data snapshot

Create a short **Data Snapshot** containing:

1. the number of countries in the dataset;
2. the mean and median of one variable of your choice;
3. one histogram of that variable;
4. **80–120 words** explaining what the distribution looks like and whether the mean or median is more informative.

Render the notebook to HTML and submit the HTML file.

> **Before submitting:** check that your code runs from top to bottom and that your figure is visible.
