---
title: "Lab 00 — Welcome to R"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
source("lab_submission.R")
un <- read_csv("data/UNpop.csv", show_col_types = FALSE) %>%
  mutate(pop_millions = world.pop / 1000)
lab_meta <- list(id = "lab00", title = "Welcome to R", open = "q4")
lab_prompts <- c(
  q1 = "What does one row of this dataset represent?",
  q2 = "How many rows are in the file?",
  q3 = "Which statement fits the series?",
  q4 = "In one or two sentences, describe the series without using a causal verb."
)
```

## What you will do

Load the UN population file from QSS Chapter 1, name the unit of analysis, and read a simple summary. At the end, download a `.txt` and upload it.

```{r peek, exercise=TRUE, exercise.lines=4}
glimpse(un)
nrow(un)
```

```{r q1}
question(
  lab_prompts[["q1"]],
  answer("One year of world population", correct = TRUE),
  answer("One country"),
  answer("One variable"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

```{r q2}
question(
  lab_prompts[["q2"]],
  answer("7", correct = TRUE),
  answer("50"),
  answer("1950"),
  allow_retry = TRUE
)
```

```{r pop-summary, exercise=TRUE, exercise.lines=6}
un %>%
  summarise(
    mean_millions = mean(pop_millions),
    median_millions = median(pop_millions)
  )
```

```{r pop-plot, exercise=TRUE, exercise.lines=6}
ggplot(un, aes(x = year, y = pop_millions)) +
  geom_line() +
  geom_point() +
  labs(x = "Year", y = "World population (millions)")
```

```{r q3}
question(
  lab_prompts[["q3"]],
  answer("Population is higher in each later decade. The mean sits a little above the median.", correct = TRUE),
  answer("Population falls after 1980."),
  answer("The file is a cross-section of countries in one year."),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

```{r q4}
question_text(
  lab_prompts[["q4"]],
  open_answer(),
  placeholder = "What rises, over which years, and in what units?",
  trim = TRUE
)
```


## Submit

Type your name, download the text file, and upload it. Submit the questions first so the file contains your answers. If nothing downloads, the same `.txt` is in this folder: select it in the Files pane and choose Export.

```{r submit-ui, echo=FALSE}
shiny::textInput("student_name", "Your name", placeholder = "Name Surname")
shiny::downloadButton("download_txt", "Download submission .txt")
```

```{r submit-server, context="server"}
output$download_txt <- shiny::downloadHandler(
  filename = function() lab_submission_filename(input$student_name, lab_meta$id),
  content = function(file) {
    write_lab_submission(
      path = file,
      name = input$student_name,
      lab_id = lab_meta$id,
      lab_title = lab_meta$title,
      state = learnr::get_tutorial_state(),
      prompts = lab_prompts,
      open = lab_meta$open
    )
    # The viewer in Posit Cloud often blocks the browser download.
    file.copy(
      file,
      lab_submission_filename(input$student_name, lab_meta$id),
      overwrite = TRUE
    )
  }
)
```

