---
title: "Lab 07 — Randomized Experiments"
subtitle: "Sessions 16–17: Randomized Experiments, Parts 1–2"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---


## Learning goals

You will analyze a randomized experiment, check treatment balance, estimate a treatment effect, and explain why randomization supports causal inference.

## 1. Setup

```{r}
library(tidyverse)
exp <- read_csv("data/diplomacy_experiment.csv")
```

The data are synthetic. Participants were randomly assigned to either:

- `treatment_message = 1`: a diplomatic-cooperation message;
- `treatment_message = 0`: a control message.

Outcome: support for diplomatic cooperation after treatment.

## 2. Check treatment allocation

```{r}
exp |>
  count(treatment_message)
```

## 3. Check balance

Randomization should make pretreatment variables similar *on average*.

```{r}
exp |>
  group_by(treatment_message) |>
  summarise(
    mean_age = mean(age),
    mean_prior_support = mean(prior_support),
    n = n()
  )
```

Are the groups exactly identical? Do they need to be?

## 4. Difference in means

```{r}
exp |>
  group_by(treatment_message) |>
  summarise(
    mean_post_support = mean(post_support_diplomatic_cooperation),
    n = n()
  )
```

Calculate the difference manually.

## 5. Regression estimate

```{r}
m_exp <- lm(
  post_support_diplomatic_cooperation ~ treatment_message,
  data = exp
)

summary(m_exp)
confint(m_exp)
```

How does the treatment coefficient relate to the difference in means?

## 6. Optional precision adjustment

```{r}
m_exp_adj <- lm(
  post_support_diplomatic_cooperation ~ treatment_message + prior_support,
  data = exp
)

summary(m_exp_adj)
```

Why is `prior_support` not needed to *create* causal identification under random assignment?

## DELIVERABLE — Experimental Results Brief

Write a **200–250 word brief** for a policymaker.

Include:

- treatment and control means;
- estimated average treatment effect;
- 95% confidence interval;
- a one-sentence statistical conclusion;
- a one-sentence substantive conclusion;
- a paragraph answering: **Why is the causal interpretation stronger here than in an observational correlation?**

Submit the rendered HTML.
