---
title: "Lab 02 — Correlation"
subtitle: "Sessions 4–5: Correlation, Parts 1–2"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---


## Learning goals

You will investigate association between two variables using scatterplots and correlation, and practice separating **association** from **causal interpretation**.

## 1. Setup

```{r}
library(tidyverse)
world <- read_csv("data/country_indicators.csv")
```

## 2. Visual association

Question:

> Are richer countries more democratic?

```{r}
ggplot(world, aes(x = gdp_per_capita, y = democracy_score)) +
  geom_point() +
  labs(
    title = "GDP per capita and democracy",
    x = "GDP per capita",
    y = "Democracy score"
  )
```

GDP per capita is strongly right-skewed. Try the log scale:

```{r}
ggplot(world, aes(x = log(gdp_per_capita), y = democracy_score)) +
  geom_point() +
  labs(
    title = "Economic development and democracy",
    x = "Log GDP per capita",
    y = "Democracy score"
  )
```

Which plot is easier to interpret?

## 3. Correlation coefficient

```{r}
cor(log(world$gdp_per_capita), world$democracy_score)
```

Interpret the **sign** and approximate **strength** of the association.

## 4. Correlation is not a causal effect

Complete the sentence:

> A positive correlation between GDP and democracy does not by itself show that increasing GDP would cause democracy to increase because ...

Give at least two possible explanations.

## 5. Explore another relationship

Choose one pair:

- democracy and life expectancy;
- democracy and military spending;
- trade openness and GDP;
- UN voting alignment and democracy.

Create a scatterplot and calculate the correlation.

```{r}
# Your code here
```

## 6. Spot the outliers

Use the plot to identify observations that do not follow the overall pattern.

```{r}
world |>
  select(country, gdp_per_capita, democracy_score) |>
  arrange(desc(gdp_per_capita))
```

Do unusual cases weaken, strengthen, or complicate your substantive argument?

## DELIVERABLE — Association Brief

Produce a **150–220 word Association Brief** answering one international-relations question.

Include:

- one labelled scatterplot;
- the correlation coefficient;
- a plain-language interpretation of direction and strength;
- one observation about an unusual country/case;
- a final paragraph headed **“Causal interpretation”** explaining why the result does or does not support a causal claim.

Submit the rendered HTML.
