---
title: "Lab 02 — Correlation"
subtitle: "Sessions 4–5: Correlation, Parts 1–2"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---

## Learning goals

You will plot two variables and read a correlation as association. The file is the congressional roll-call data in QSS 3.6. `dwnom1` and `dwnom2` are two DW-NOMINATE scores for a member of Congress. They are measures of voting patterns, not a treatment and its outcome.

## 1. Setup

```{r}
library(tidyverse)
congress <- read_csv("data/congress.csv")
congress112 <- congress %>%
  filter(congress == 112, party %in% c("Democrat", "Republican"))
```

## 2. The scatter and the correlation

```{r}
ggplot(congress112, aes(x = dwnom1, y = dwnom2, color = party)) +
  geom_point(alpha = 0.7) +
  labs(
    x = "DW-NOMINATE dimension 1",
    y = "DW-NOMINATE dimension 2",
    title = "112th Congress"
  )
```

```{r}
cor(congress112$dwnom1, congress112$dwnom2)
```

The correlation is about 0.12. In this Congress the two scores move together only weakly. A positive correlation says that higher values on one score tend to come with higher values on the other. It does not say that one score causes the other, and it does not say that a party caused a member's voting record.

## Source

`congress.csv` is taken unchanged from [kosukeimai/qss](https://github.com/kosukeimai/qss) `MEASUREMENT/congress.csv`. Sessions 4–5 in [mir-prob-stats-QQS](https://github.com/iedaejin/mir-prob-stats-QQS). See `data/SOURCE.txt`.
