---
title: "Lab 05 — Estimation, Uncertainty and Hypothesis Testing"
subtitle: "Sessions 11–12: Estimation and Uncertainty; Hypothesis Testing"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---


## Learning goals

You will connect random sampling to sampling distributions, confidence intervals, and hypothesis tests.

## 1. Setup

```{r}
library(tidyverse)
survey <- read_csv("data/survey_attitudes.csv")
```

Treat this dataset as a teaching population.

## 2. Take one random sample

```{r}
set.seed(123)
samp <- survey |>
  slice_sample(n = 40)

mean(samp$support_international_cooperation)
```

How close is it to the population mean?

```{r}
mean(survey$support_international_cooperation)
```

## 3. Build a sampling distribution

```{r}
sample_means <- replicate(
  1000,
  mean(sample(
    survey$support_international_cooperation,
    size = 40,
    replace = TRUE
  ))
)

tibble(sample_mean = sample_means) |>
  ggplot(aes(x = sample_mean)) +
  geom_histogram(bins = 30) +
  labs(
    title = "Sampling distribution of the mean",
    x = "Sample mean",
    y = "Frequency"
  )
```

What does this distribution represent?

## 4. Confidence interval

```{r}
t.test(samp$support_international_cooperation)$conf.int
```

Interpret the interval in words.

## 5. Hypothesis testing

Question:

> Do postgraduate respondents report different average support for international cooperation than respondents with a Bachelor degree?

```{r}
two_groups <- survey |>
  filter(education %in% c("Bachelor", "Postgraduate"))

t.test(
  support_international_cooperation ~ education,
  data = two_groups
)
```

Identify:

- group means;
- estimated difference;
- p-value;
- confidence interval.

## 6. Statistical vs substantive significance

A small p-value does not automatically imply a substantively important difference.

Write two separate conclusions:

**Statistical conclusion:**  
...

**Substantive conclusion:**  
...

## DELIVERABLE — Inference Memo

Submit an **Inference Memo** that answers one group-comparison question using the survey data.

It must include:

- the two group means;
- the estimated difference;
- a 95% confidence interval;
- the p-value;
- a **statistical conclusion**;
- a **substantive conclusion**;
- **150–220 words** total.

Submit the rendered HTML.
