---
title: "Lab 05 — Estimation, Uncertainty and Hypothesis Testing"
subtitle: "Sessions 11–12: Estimation and Uncertainty; Hypothesis Testing"
format:
  html:
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---

## Learning goals

You will read a difference in means, a confidence interval, and a p-value. The file is the STAR class-size experiment used in QSS 7.1. `classtype` 1, 2, and 3 are small class, regular class, and regular class with an aide. The outcome here is the fourth-grade reading score, `g4reading`.

## 1. Setup

```{r}
library(tidyverse)
star <- read_csv("data/STAR.csv") %>%
  mutate(classtype = factor(
    classtype,
    labels = c("Small class", "Regular class", "Regular class with aide")
  )) %>%
  filter(
    !is.na(g4reading),
    classtype %in% c("Small class", "Regular class")
  )
```

```{r}
star %>%
  group_by(classtype) %>%
  summarise(mean_reading = mean(g4reading), n = n(), .groups = "drop")
```

```{r}
t.test(g4reading ~ classtype, data = star)
```

The small-class mean is about 723. The regular-class mean is about 720. The difference is a few points. The 95 percent interval runs from about −1.7 to 8.7 and includes zero. The two-sided p-value is about 0.19.

A 95 percent interval is a procedure: in repeated samples, 95 percent of such intervals would cover the true difference. It is not a 95 percent probability that this particular interval contains the truth. A p-value is the probability of a gap at least this large if the two class types had the same mean. Failing to reject is not proof that class size does not matter. The estimate is small and imprecise.

Random assignment is what makes this comparison a candidate for an effect of class type. The interval does not create that design.

## Source

`STAR.csv` is taken unchanged from [kosukeimai/qss](https://github.com/kosukeimai/qss) `UNCERTAINTY/STAR.csv`. Sessions 11–12 in [mir-prob-stats-QQS](https://github.com/iedaejin/mir-prob-stats-QQS). See `data/SOURCE.txt`.
